---
title: "Logistic Regression Project"
author: "Team 18"
date: "8/27/2024"
output: html_document
---

# Logistic Regression Project

## Phase #1

```{r}
# Load required packages
library(readr)
library(tidyverse)
# library(DataExplorer)
library(mgcv)
library(vcdExtra)
```

### Loading Data

This set represents insurance training data. This is the only dataset to be used for the first Phase.

```{r}
# Read data

insurance_t <- read_csv("insurance_t.csv")
head(insurance_t)
```

### Explore the data

```{r}
tibble::glimpse(insurance_t)
```

### Assessing Missing Values

```{r}
## Pull a dataframe of all missing value counts
missing_data <- data.frame(
  col_name=names(colSums(is.na(insurance_t))),
  missing_count=colSums(is.na(insurance_t)),
  missing_perc=(colSums(is.na(insurance_t))/dim(insurance_t)[[1]])*100
  )
# Convert to tibble
missing_data <- as.tibble(missing_data)
## Pull the count and percentage of missing values for all variables that have them
missing_data <- missing_data %>% 
  filter(missing_count!=0) %>% 
  arrange(desc(missing_count))
missing_data
```

```{r}
# Bar plot of % values missing in columns

highest_missing_order = missing_data$col_name
ggplot(missing_data, aes(
#  x=col_name, 
  x=factor(col_name, level = highest_missing_order),
  y=missing_perc)) + 
  geom_bar(stat = "identity",fill="#4DD0E1") + 
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5)) +
  xlab('Column') + 
  ylab('% records missing') + 
  ggtitle('Missing values by column')
```

```{r}
# EDA

# create_report(insurance_t)
```

Import Reference Sheet for Variable Types

```{r}
# Load in a csv that has all predictor variables and their corresponding "class" (ordinal, continuous, nominal, binary). This was created manually.

var_info <- read_csv("Logistic_variable_classes.csv")
```

### Explore Predictor Variables

Binary Target Variable (INS) for whether or not customer bought the insurance product.

```{r}
# Convert categorical variables to factor based on class determined above

for (var_row in 1:nrow(var_info)) {
  
  # Extract the variable name and its class
  var <- var_info[[var_row, 1]]
  class <- var_info[[var_row, 2]]
  
  # Check if the variable class is one that needs to be factored
  if (class %in% c("Binary", "Nominal", "Ordinal")) {
    
    # Convert the column to factor in the insurance_t dataframe
    insurance_t[[var]] <- factor(insurance_t[[var]])
  }
}
## Convert the target variable to factor 
insurance_t$INS <- factor(insurance_t$INS)
```

Build table of p-values ranked by significance

```{r}
# Loop through each predictor variable
for (var_row in 1:nrow(var_info)) {
    
    var <- var_info[[var_row, 1]]
    class <- var_info[[var_row, 2]]
    
    if (class == "Continuous") {
      # Logistic regression with the continuous variable
      logit.model.var <- glm(INS ~ insurance_t[[var]], data = insurance_t, family = binomial(link = "logit"))
      var_info$p.value[var_row] <- summary(logit.model.var)$coefficients[, "Pr(>|z|)"][[2]][1]
    }
    # Mantel-Haenszel Chi-Square Test if Binary or Ordinal
    else if (class == "Binary" | class == "Ordinal") {
      cmhtest_vals <- CMHtest(table(insurance_t$INS, insurance_t[[var]]))
      var_info$p.value[var_row] <- cmhtest_vals$table[1, ][[3]][1]
    }
    
    # Pearson Chi square test for nominal variables
    else {
      var_info$p.value[var_row] <- chisq.test(table(insurance_t$INS, insurance_t[[var]]))[[3]][1]
    }
}
  
# Create a table of pvalues that have been pulled of only the significant variables
filtered_var_table <- var_info %>%
  mutate(p.value = as.numeric(p.value)) %>%
  filter(p.value <= .0002) %>%
  arrange(p.value)
    
print(filtered_var_table)
```

### Odds Ratio Table

```{r}
## Collect all binary predictor variables
binary_vars <- var_info %>%
  filter(Class == "Binary") %>%
  pull(Variable)

## Convert to tibble
binary_odds_ratios <- as_tibble(binary_vars)
#binary_odds_ratios$odds_ratio <- NA

## Loop through each variable and pull the odds ratio
cnt = 1
for (var in binary_vars) {
  
  ## Create binary logistic regression for binary variables
  logit.model.binary <- glm(INS ~ insurance_t[[var]], data = insurance_t, family = binomial(link = "logit"))
  
  ## Calculate odds ratio
  binary_odds_ratios$odds_ratio[cnt] <- exp(coef(logit.model.binary))[2]
  cnt <- cnt + 1
  
}

## Arrange the odds ratio in descending order
binary_odds_ratios <- binary_odds_ratios %>%
  arrange(desc(odds_ratio))

print(binary_odds_ratios)
```

-   For aspects that reflect a challenging or perhaps even low-income financial environment, these elements such as having taken out a loan (ILS), having insufficient funds (NSF), or even ATM transactions (ATM) correspond with a lesser likelihood of taking out a loan

-   Elements that reflect activities with long-term investing such as having a money market (MM), IRA (IRA), certificate of deposit (CD), or investment account (INV) reflect a greater likelihood to purchase a variable rate annuity product.

-   A customer has a 3.47x greater odds of purchasing a variable rate annuity product if they have an investment account compared to if they do not.

### Assessing Linearity

```{r}
## Collect all continuous variables
continuous_vars <- var_info %>%
  filter(Class == "Continuous") %>%
  pull(Variable)

## Convert to tibble
continuous_linearity <- as_tibble(continuous_vars)

## Loop through all continuous variables and test the linearity assumption
cnt = 1
for (var in continuous_vars) {
  
  subset <- insurance_t[,c("INS", var)]
  
  ## Fit a normal logistci regression function for each variable
  logit.model.contin <- glm(INS ~ ., data = subset, family = binomial(link = "logit"))
  ## Give an s function for just the continuous variables. This tells the   GAM function to calculate a spline
  fit.gam <- mgcv::gam(reformulate(paste('s(',var,')'), 'INS'),
   data = subset, family = binomial(link = 'logit'), method = 'REML')
  
  ## Check via hypothesis test if the models are statistically different. 
  contin_lin_test <- anova(logit.model.contin, fit.gam, test = 'LRT')
  
  ## Classify if the variable meets the linearity assumption
  continuous_linearity$linearity[cnt] <- ifelse(contin_lin_test$`Pr(>Chi)`[2] <= .002, 'No', 'Yes')
  ## Pull the p-value
  continuous_linearity$p.val[cnt] <- contin_lin_test$`Pr(>Chi)`[2]
  cnt <- cnt + 1
  
}

## Arrange based oon descending order of p-values
continuous_linearity <- continuous_linearity  %>%
  arrange(desc(p.val))

print(continuous_linearity)
```
